---
title: "Understanding patterns of species abundance and persistence in TROLL"
output: 
    html_document:
        toc: true
        toc_float: true
        number_section: true
        highlight: zenburn
knit: (function(inputFile, encoding) 
        {
          rmarkdown::render(inputFile, encoding = encoding, output_dir = "markdowns") 
        }
      )
---

```{r setup, include=FALSE}

options(width = 1000)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 8,
                      fig.width = 12) 

library(data.table)
library(tidyverse)
library(ggthemes)
library(viridis) 
library(cowplot)
library(scales)
library(plotly)
library(randomForest)

set.seed(1)

```

```{r load-data, echo=FALSE, message=FALSE, warning=FALSE}

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("../../data/raw/")

troll <- readRDS("TROLL_Table1.rds") %>%
    select(-LineNumber, -Productivity) %>%
    mutate(SpeciesID = as.factor(SpeciesID),
           SpeciesID = as.numeric(SpeciesID))

troll_traits <- fread("TROLL_Table2.txt") %>%
    rename(SpeciesID = species_binomial) %>%
    mutate(SpeciesID = as.factor(SpeciesID),
           SpeciesID = as.numeric(SpeciesID))

troll <- troll %>%
    mutate(Model = as.character(Model),
           Ninitial = as.factor(Ninitial),
           Rep = as.numeric(Rep),
           SeedRain = as.factor(SeedRain),
           SpeciesID = as.character(SpeciesID),
           Stage = as.factor(Stage),
           Year = as.numeric(Year),
           Biomass = as.numeric(Biomass))

troll <- troll %>%
    mutate(Stage = recode(Stage,
                          assembly = "metacommunity",
                          disassembly = "isolation"))

troll <- troll %>%
    filter(SeedRain %in% c(100))

troll <- troll %>%
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))

troll <- troll %>%
    select(-Model, -SeedRain) %>%
    mutate(SpeciesID = as.factor(SpeciesID))

troll_traits <- troll_traits %>%
    mutate(SpeciesID = as.factor(SpeciesID))

troll <- inner_join(troll, troll_traits, by = c("SpeciesID"))

troll <- troll %>%
    mutate(h_realmax = hmax * dmax / (dmax + ah)) %>%
    select(-hmax, -ah)

```

In this document, I examine how a species' performance in mixture depends on 1) its monoculture biomass
and 2) its trait values. To measure performance in mixture, I use two indices: 1) total biomass and 2)
time to extinction.

# How trait values translate into monoculture biomass

First, a facetted plot of the relationship between each species' trait values and its corresponding
monoculture biomass. 

```{r monoculture-biomass-vs-traits, echo=FALSE, warning=FALSE, message=FALSE}

troll.monocultures <- troll %>%
    filter(Ninitial == 1,
           Year %in% c(100, 200)) %>%
    rename(monoculture.biomass = Biomass) %>%
    select(-Ninitial, -Year, -Rep) %>%
    pivot_longer(c(-SpeciesID, -Stage, -monoculture.biomass),
                 names_to = "trait", 
                 values_to = "trait.value")

troll.monocultures <- troll.monocultures %>%
    group_by(trait) %>%
    nest() %>%
    mutate(plot = map2(.x = data,
                       .y = trait,
                       .f = ~ {
                           ggplot(.x) +
                               geom_point(aes(x = trait.value,
                                              y = monoculture.biomass,
                                              color = SpeciesID),
                                          size = 3,
                                          show.legend = FALSE) +
                               facet_grid(cols = vars(Stage)) +
                               labs(x = "Trait value",
                                    y = "Monoculture biomass",
                                    title = .y) +
                               theme_bw()
                       }))

plot_grid(plotlist = troll.monocultures$plot,
          ncol = 2)

```

Color indicates SpeciesID. From visual inspection, it appears that `dmax` and `h_realmax` are the most important traits 
determining monoculture biomass.

Because I'm going to be plotting the species' traits in relation to each other and their species' emergent biomass,
I need to pick which traits are the most important. Therefore, I use a random forest to find the importance for 
each of the traits in predicting monoculture biomass. Note that, to increase the amount of data for the random forest,
I use the last 20 years of the simulation as input data for the random forest, adding `Year` as a predictor 
to ensure that temporal autocorrelation does not bias the variable importance measures (see
[here](https://www.researchgate.net/post/Does_random_forest_account_for_temporal_auto-correlation)).

```{r monoculture-biomass-vs-traits-randomForest-varImp, echo=TRUE, warning=FALSE, message=FALSE}

troll.monocultures <- troll %>%
    filter(Ninitial == 1,
           Year %in% seq(180, 200)) %>%
    rename(monoculture.biomass = Biomass) %>%
    mutate(id = row_number()) %>%
    mutate_if(is.character, as.factor) %>%
    select(-SpeciesID, -Ninitial, -Stage, -Rep) 

train <- troll.monocultures %>% sample_frac(.70)
test <- anti_join(troll.monocultures, train, by = 'id')

train <- train %>% select(-id)
test <- test %>% select(-id)

rf <- randomForest(data = as.data.frame(train), 
                   ntree = 1000, 
                   mtry = 3,
                   monoculture.biomass ~ ., 
                   importance = TRUE)

pred <- data.frame(pred = predict(rf, test))
title = paste("correlation: ", round(cor(pred, test$monoculture.biomass)[[1]], 2),
              "  |  mean squared error: ", round(mean(rf$mse), 2),
              "  |  R-squared: ", round(mean(rf$rsq), 2),
              sep = "")

varImpPlot(rf, main = title)

```

Strangely, incorporating `h_realmax`, rather than its components `ah` and `hmax` leads to the the year becoming
an important contributing variable (in general, one should pay more attention to `%IncMSE` than `IncNodePurity`). This
was not the case in my previous document. To get a handle on it, I'm going to first plot the species biomasses over
time.

```{r monoculture-biomass-over-time, echo=TRUE, warning=FALSE, message=FALSE}

monocultures <- troll %>%
    filter(Ninitial == 1,
           Year %in% seq(180, 200)) %>%
    rename(monoculture.biomass = Biomass) %>%
    select(-Ninitial, -Stage, -Rep) 

ggplot(monocultures) +
    geom_line(aes(x = Year,
                  y = monoculture.biomass,
                  color = as.factor(SpeciesID)),
              show.legend = FALSE) +
    labs(y = "Monoculture biomass") +
    theme_bw() +
    theme(aspect.ratio = 1)

```

So, there is definitely still one species that's growing until the 200th timestep. I wonder if 
the reason for the original random forest's confusion is that the composite variable `h_realmax` is more 
unique to a single species than the other three variables its composed of were. Therefore, the fact that
this one species is still growing becomes signal rather than noise. I think that this is especially the case
because `dmax`, which was once the most predictive variable, is now included in the dataset twice: 
Once by itself, and again as a component in `h_realmax`.

Given that `Year` is now an important predictor, I'm going to run the randomForest again except only
for the terminal year.

```{r monoculture-biomass-vs-traits-randomForest-varImp-Again, echo=TRUE, warning=FALSE, message=FALSE}

troll.monocultures <- troll %>%
    filter(Ninitial == 1,
           Year == 200) %>%
    rename(monoculture.biomass = Biomass) %>%
    mutate(id = row_number()) %>%
    mutate_if(is.character, as.factor) %>%
    select(-SpeciesID, -Ninitial, -Stage, -Rep) 

train <- troll.monocultures %>% sample_frac(.70)
test <- anti_join(troll.monocultures, train, by = 'id')

train <- train %>% select(-id)
test <- test %>% select(-id)

rf <- randomForest(data = as.data.frame(train),
                   ntree = 1000,
                   mtry = 3,
                   monoculture.biomass ~ .,
                   importance = TRUE)

pred <- data.frame(pred = predict(rf, test))
title = paste("correlation: ", round(cor(pred, test$monoculture.biomass)[[1]], 2),
              "  |  mean squared error: ", round(mean(rf$mse), 2),
              "  |  R-squared: ", round(mean(rf$rsq), 2),
              sep = "")

varImpPlot(rf, main = title)

```

The correlation is less, as is the R-squared, but we're getting a better idea of what's actually important
in terms of traits. Therefore, the new plots will use `h_realmax` and `dmax` as their focal traits.

I want to note that this is not necessarily good practice: The training dataset has only 45 data points in it, 
which is far too few for a random forest. I can spend some more time figuring out how to do this better, but as a 
first pass it confirms my intuition so I'll leave it. You can tell me if you want something more robust.

```{r monoculture-biomass-vs-traits-specific, echo=TRUE, warning=FALSE, message=FALSE}

troll.monocultures <- troll %>%
    filter(Ninitial == 1,
           Year %in% c(100, 200)) %>%
    select(-Ninitial, -Year, -Rep)

ggplot(troll.monocultures) +
    geom_point(aes(x = dmax, 
                   y = h_realmax,
                   color = Biomass),
               size = 4) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis() +
    labs(color = "Monoculture\nbiomass") +
    theme_bw()

```  

****************************************************************************************************

# Competitive ability in mixture vs. monoculture biomass

I now explore how the competitiveness of a species in a community (either 32- or, later, 64-species)
depends on 1) its monoculture biomass, and subsequently 2) its trait values.

## 32-species mixtures

### Species' biomass in 32-species mixture vs. its biomass in monoculture

```{r 32-species-biomass-vs-monoculture-biomass, echo=TRUE, warning=FALSE, message=FALSE}

troll.monocultures <- troll %>%
    filter(Ninitial == 1,
           Year %in% c(100, 200)) %>%
    rename(monoculture.biomass = Biomass) %>%
    select(-Ninitial, -Year, -Rep) 

troll.32 <- troll %>%
    filter(Ninitial == 32,
           Year %in% c(100, 200)) %>%
    select(SpeciesID, Stage, Biomass, dmax, wsg, lma, pmass, nmass, h_realmax)

troll.combined <- inner_join(troll.monocultures, troll.32)

ggplot(troll.combined) +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID)) +
    facet_grid(cols = vars(Stage)) +
    labs(x = "Monoculture biomass",
         y = "Biomass in 32-species mixture") +
    theme_bw() + 
    theme(legend.position = "none")

```

Each color represents a species, while the number of discrete points per species depends on
the number of replicates the species was included in. Inspecting this relationship visually, it appears 
that species with low biomass monocultures stand a better chance to dominate 32-species mixtures. 
However, many relatively high biomass monocultures remain in the community.

### Biomass in 32-species mixture vs. trait values

To understand this relationship more deeply, I inspect how species' traits impact their performance 
in 32-species mixtures.

```{r 32-species-biomass-vs-traits, echo=TRUE, warning=FALSE, message=FALSE}

troll.32 <- troll %>%
    filter(Ninitial == 32,
           Year %in% c(100, 200)) %>%
    select(Stage, SpeciesID, Biomass, dmax, wsg, lma, pmass, nmass, h_realmax)

troll.32 <- troll.32 %>%
    pivot_longer(c(-Stage, -SpeciesID, -Biomass),
                 names_to = "trait",
                 values_to = "trait.value")

troll.32 <- troll.32 %>%
    group_by(trait) %>%
    nest() %>%
    mutate(plot = map2(.x = data,
                       .y = trait,
                       .f = ~ {
                           ggplot(.x) +
                               geom_point(aes(x = trait.value,
                                              y = Biomass,
                                              color = SpeciesID),
                                          size = 1,
                                          show.legend = FALSE) +
                               facet_grid(cols = vars(Stage)) +
                               labs(x = "Trait value",
                                    y = "Monoculture biomass",
                                    title = .y) +
                               theme_bw()
                       }))

plot_grid(plotlist = troll.32$plot,
          ncol = 2)
```

It is difficult to discern which trait is the most important predictor of total biomass within the
32-species mixtures. Therefore, I again employ a randomForest to find the traits' importances.

To increase the data available to the random forest, I again use the last 20 years of the dataset 
to generate the random forest model, incorporating `Year` as a predictor to ensure that temporal 
autocorrelation plays a minimal role in biasing the results.

```{r 32-species-biomass-vs-traits-randomForest-varImp-isolation, echo=TRUE, warning=FALSE, message=FALSE}

troll.32 <- troll %>%
    filter(Ninitial == 32,
           Year %in% seq(180, 200))

troll.32 <- troll.32 %>%
    ungroup() %>%
    mutate(id = row_number()) %>%
    mutate_if(is.character, as.factor) %>%
    select(-Ninitial, -Rep, -Stage, -SpeciesID) 

train <- troll.32 %>% sample_frac(.70)
test <- anti_join(troll.32, train, by = 'id')

train <- train %>% select(-id)
test <- test %>% select(-id)

rf <- randomForest(data = as.data.frame(train),
                   Biomass ~ .,
                   importance = TRUE)

pred <- data.frame(pred = predict(rf, test))

title = paste("correlation: ", round(cor(pred, test$Biomass)[[1]], 2),
              "  |  mean squared error: ", round(mean(rf$mse), 2),
              "  |  R-squared: ", round(mean(rf$rsq), 2),
              sep = "")

varImpPlot(rf, main = title)

``` 

`lma` is the best predictor of total biomass in mixture, though all traits
play an important roll in determining a species' biomass within a community. `Year` is a poor predictor 
of species' biomass in monoculture, indicating that sampling the last 20 years doesn't skew the importance values.

```{r 32-species-biomass-vs-traits-3D-metacommunity-isolation, echo=TRUE, warning=FALSE, message=FALSE}

troll.32 <- troll %>%
    filter(Ninitial == 32,
           Year == 200) %>%
    group_by(SpeciesID, dmax, wsg, lma, pmass, nmass, h_realmax) %>%
    summarise(Biomass = mean(Biomass)) %>%
    filter(Biomass > 0)

plot_ly(troll.32,
        x = ~wsg, 
        y = ~pmass,
        z = ~lma,
        color = ~log(Biomass),
        mode = "markers",
        type = "scatter3d")

```

This randomForest was purely for the isolation stage, so I'm including one below that belongs to the
metacommunity stage. From our experience so far, I would expect that the traits important for competitiveness
during the metacommunity stage are different from those during the isolation stage.

```{r 32-species-biomass-vs-traits-randomForest-varImp-metacommunity, echo=TRUE, warning=FALSE, message=FALSE}

troll.32 <- troll %>%
    filter(Ninitial == 32,
           Year %in% seq(80, 100))

troll.32 <- troll.32 %>%
    ungroup() %>%
    mutate(id = row_number()) %>%
    mutate_if(is.character, as.factor) %>%
    select(-Ninitial, -Rep, -Stage, -SpeciesID) 

train <- troll.32 %>% sample_frac(.70)
test <- anti_join(troll.32, train, by = 'id')

train <- train %>% select(-id)
test <- test %>% select(-id)

rf <- randomForest(data = as.data.frame(train),
                   Biomass ~ .,
                   importance = TRUE)

pred <- data.frame(pred = predict(rf, test))

title = paste("correlation: ", round(cor(pred, test$Biomass)[[1]], 2),
              "  |  mean squared error: ", round(mean(rf$mse), 2),
              "  |  R-squared: ", round(mean(rf$rsq), 2),
              sep = "")

varImpPlot(rf, main = title)

``` 

The central difference, it seems, is that `h_realmax` and `dmax` are significantly more important to predicting
the total biomass in mixture during the metacommunity stage, than during the isolation phase. I would conjecture that
this is because those two traits are very important in determining monoculture biomass, which is in turn used to 
calculate seed addition. **This makes me wonder, Isabelle, when you implemented the SimNet protocol into TROLL, 
did you have the amount of seed addition (i.e. kg/species of seeds) be equal to the average seed addition across all
monocultures, or across their own individual monocultures?** This would better explain, perhaps, why the higher 
`h_realmax` and `dmax` species seem to benefit more seed addition than the others. But maybe this is true even without 
any mistakes in the implementation, because species with higher wood density are hearty but slow growing, so
the seed addition is more beneficial to them?

```{r 32-species-biomass-vs-traits-3D-metacommunity, echo=TRUE, warning=FALSE, message=FALSE}

troll.32 <- troll %>%
    filter(Ninitial == 32,
           Year == 100) %>%
    group_by(SpeciesID, dmax, wsg, lma, pmass, nmass, h_realmax) %>%
    summarise(Biomass = mean(Biomass)) %>%
    filter(Biomass > 0)

plot_ly(troll.32,
        x = ~h_realmax, 
        y = ~dmax,
        z = ~lma,
        color = ~log(Biomass),
        mode = "markers",
        type = "scatter3d")

```
