---
title: "Pearson's correlations"
output: 
    html_document:
        toc: true
        toc_float: true
        number_section: true
        fig_width: 7
        fig_height: 4
knit: (function(inputFile, encoding) 
        {
          rmarkdown::render(inputFile, encoding = encoding, output_dir = "markdowns") 
        }
      )
---

<style>
pre 
{
    overflow-x: auto;
}
pre code 
{
    word-wrap: normal;
    white-space: pre;
}
</style>

```{r setup, include=FALSE}

options(width = 1000)

library(tidyverse)

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      size = "small",
                      dev = "png") 

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
source("SimNet_ReadModels.R")

models <- model_runs %>%
    bind_rows() %>%
    select(-SeedRain, -Year, -Productivity)

models.mono <- models %>%
    filter(Ninitial == 1) %>%
    rename(monoculture.biomass = Biomass) %>%
    ungroup() %>%
    select(-Ninitial, -Rep)

models <- inner_join(models, models.mono) %>%
    filter(Ninitial != 1,
           Model != "IBC_grass.noNDD")

models$Ninitial <- droplevels(models$Ninitial)

```

This document describes the Pearson's correlation within each model between the species' monoculture biomasses and their biomasses within the 32-species mixtures during both seed addition phases. 


****************************************************************************************************
# Grass1

<i> Clark, A.T., C. Lehman, and D. Tilman. 2018. Identifying mechanisms that structure ecological communities by snapping model parameters to empirically observed trade-offs. Ecology Letters 21:494–505. </i>

```{r Grass1-setup, include=FALSE}

Grass1 <- models %>%
    filter(Model == "adam") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Grass1-plot-64}

ggplot(Grass1 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass1-select-32, include=FALSE}

Grass1 <- Grass1 %>%
    filter(Ninitial == 32)

```

## 32-species mixture

We now plot the relationship between a species' monoculture biomass and its biomass within the 32-species mixture dataset to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within the richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Grass1-plot}

ggplot(Grass1)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass1-correlation}

Grass1.corr <- Grass1 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Grass1.corr <- Grass1.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Grass1.corr, 
             caption = "Correlation between monoculture and mixture biomass",
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```


****************************************************************************************************
# Grass2

<i> Turnbull, L. A., J. M. Levine, M. Loreau, and A. Hector. 2013. Coexistence, niches and biodiversity effects on ecosystem functioning. Ecology Letters 16:116–127. </i> 

```{r Grass2-setup, include=FALSE}

Grass2 <- models %>%
    filter(Model == "lindsay") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Grass2-plot-64}

ggplot(Grass2 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass2-select-32, include=FALSE}

Grass2 <- Grass2 %>%
    filter(Ninitial == 32)

```

## 32-species mixture

We now plot the relationship between a species' monoculture biomass and its biomass within the 32-species mixture dataset to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within the richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Grass2-plot}

ggplot(Grass2)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass2-correlation}

Grass2.corr <- Grass2 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Grass2.corr <- Grass2.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Grass2.corr, 
             caption = "Correlation between monoculture and mixture biomass",
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```


****************************************************************************************************
# Grass3

<i> May, F., V. Grimm, and F. Jeltsch. 2009. Reversed effects of grazing on plant diversity: The role of below-ground competition and size symmetry. Oikos 118:1830–1843. </i> 

```{r Grass3-setup, include=FALSE}

Grass3 <- models %>%
    filter(Model == "IBC_grass.NDD") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Grass3-plot-64}

ggplot(Grass3 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass3-select-32, include=FALSE}

Grass3 <- Grass3 %>%
    filter(Ninitial == 32)

```

## 32-species mixture

We now plot the relationship between a species' monoculture biomass and its biomass within the 32-species mixture dataset to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within the richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Grass3-plot}

ggplot(Grass3)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass3-correlation}

Grass3.corr <- Grass3 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Grass3.corr <- Grass3.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Grass3.corr, 
             caption = "Correlation between monoculture and mixture biomass",
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```


****************************************************************************************************
# Forest1

<i> Rüger, N., R. Condit, D. H. Dent, S. J. DeWalt, S. P. Hubbell, J. W. Lichstein, O. R. Lopez, C. Wirth, and C. E. Farrior. 2020. Demographic trade-offs predict tropical forest dynamics. Science 368:165–168. </i> 

```{r Forest1-setup, include=FALSE}

Forest1 <- models %>%
    filter(Model == "PPA") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Forest1-plot-64}

ggplot(Forest1 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Forest1-select-32, include=FALSE}

Forest1 <- Forest1 %>%
    filter(Ninitial == 32)

```

## 32-species mixture

We now plot the relationship between a species' monoculture biomass and its biomass within the 32-species mixture dataset to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within the richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Forest1-plot}

ggplot(Forest1)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Forest1-correlation}

Forest1.corr <- Forest1 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Forest1.corr <- Forest1.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Forest1.corr, 
             caption = "Correlation between monoculture and mixture biomass",
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```


****************************************************************************************************
# Forest2

<i> Maréchaux, I., and J. Chave. 2017. An individual-based forest model to jointly simulate carbon and tree diversity in Amazonia: description and applications. Ecological Monographs. </i> 

```{r Forest2-setup, include=FALSE}

Forest2 <- models %>%
    filter(Model == "TROLL") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Forest2-plot-64}

ggplot(Forest2 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Forest2-select-32, include=FALSE}

Forest2 <- Forest2 %>%
    filter(Ninitial == 32)

```

## 32-species mixture

We now plot the relationship between a species' monoculture biomass and its biomass within the 32-species mixture dataset to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within the richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Forest2-plot}

ggplot(Forest2)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Forest2-correlation}

Forest2.corr <- Forest2 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Forest2.corr <- Forest2.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Forest2.corr, 
             caption = "Correlation between monoculture and mixture biomass",
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```


****************************************************************************************************
# Dryland

<i> Reineking, B., M. Veste, C. Wissel, and A. Huth. 2006. Environmental variability and allocation trade-offs maintain species diversity in a process-based model of succulent plant communities. Ecological Modelling.</i> 

```{r Dryland-setup, include=FALSE}

Dryland <- models %>%
    filter(Model == "bjoern") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Dryland-plot-64}

ggplot(Dryland %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Dryland-select-32, include=FALSE}

Dryland <- Dryland %>%
    filter(Ninitial == 32)

```

## 32-species mixture

We now plot the relationship between a species' monoculture biomass and its biomass within the 32-species mixture dataset to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within the richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Dryland-plot}

ggplot(Dryland)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Dryland-correlation}

Dryland.corr <- Dryland %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Dryland.corr <- Dryland.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Dryland.corr, 
             caption = "Correlation between monoculture and mixture biomass",
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```
