---
title: "Pearson's correlations"
output: 
    html_document:
        toc: true
        toc_float: true
        number_section: true
        code_folding: hide
        highlight: zenburn
        fig_width: 8
        fig_height: 6
knit: (function(inputFile, encoding) 
        {
          rmarkdown::render(inputFile, encoding = encoding, output_dir = "markdowns") 
        }
      )
---

<style>
pre 
{
    overflow-x: auto;
}
pre code 
{
    word-wrap: normal;
    white-space: pre;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE) 

library(tidyverse)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
source("SimNet_ReadModels.R")

models <- model_runs %>%
    bind_rows() %>%
    select(-SeedRain, -Year, -Productivity)

models.mono <- models %>%
    filter(Ninitial == 1) %>%
    rename(monoculture.biomass = Biomass) %>%
    ungroup() %>%
    select(-Ninitial, -Rep)

models <- inner_join(models, models.mono) %>%
    filter(Ninitial != 1,
           Model != "IBC_grass.noNDD")

models$Ninitial <- droplevels(models$Ninitial)

```

This document describes the Pearson's correlation within each model between the species' monoculture biomasses and their biomasses within mixtures of 2, 4, 8, 16, and 32 species, during both seed addition phases. We then describe the process used to bootstrap these datasets and derive a better representation of the correlation between monoculture and mixture biomass.


****************************************************************************************************
# Grass1

<i> Clark, A.T., C. Lehman, and D. Tilman. 2018. Identifying mechanisms that structure ecological communities by snapping model parameters to empirically observed trade-offs. Ecology Letters 21:494–505. </i>

```{r Grass1-setup, include=FALSE}

Grass1 <- models %>%
    filter(Model == "adam") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Grass1-plot-64}

ggplot(Grass1 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass1-remove-64, include=FALSE}

Grass1 <- Grass1 %>%
    filter(Ninitial != 64)

```

## Monoculture vs. mixture biomass

We now plot the relationship between a species' monoculture biomass and its biomass within the mixture datasets to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within each planted species richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Grass1-plot}

ggplot(Grass1)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage),
               rows = vars(Ninitial)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

## Correlations

Within each planted species richness treatment and seed addition stage, we calculate the correlation between the species' biomasses in the mixtures and their biomasses in monoculture.

```{r Grass1-correlation, echo = TRUE}

Grass1.corr <- Grass1 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Grass1.corr <- Grass1.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Grass1.corr, 
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```

## Bootstrap

To mitigate the effect of sampling in our final analysis and generate a more reliable assessment of the average correlation between the species' biomass in monoculture and their biomass in mixture, rather than only using the 64 mixtures, we bootstrap the mixture datasets (n = 2500) and calculate the Pearson's correlation for each bootstrap.

```{r Grass1-bootstrap}

Grass1.bootstrap <- Grass1 %>%
    group_by(Ninitial, Stage) %>%
    nest() %>%
    mutate(cors = map(.x = data,
                      .f = ~
                          {
                              bootstraps <- .x %>% rsample::bootstraps(times = 2500)
                              
                              bootstraps <- bootstraps %>%
                                  mutate(r = map_dbl(.x = splits,
                                                     .f = ~ 
                                                         {
                                                             .x <- as.data.frame(.x)
                                                             cor(.x$Biomass, .x$monoculture.biomass)
                                                         }))
                          })) %>%
    unnest(cors)

ggplot(Grass1.bootstrap) +
    geom_histogram(aes(x = r)) +
    facet_grid(rows = vars(Ninitial),
               cols = vars(Stage), 
               scales = "free") +
    labs(x = "Pearson's correlation (r)",
         y = "Frequency") +
    theme_bw()

```

Based on the bootstrapping process, we now have a more complete picture of the distribution of Pearson's correlation of the species' mixture biomass to their monoculture biomass. We use this distribution of Pearson's correlations in the main analysis by relating each bootstrapped correlation to an estimate of the relationship between realized Shannon diversity and biomass, drawn from the posterior distribution of the BRMS models.


****************************************************************************************************
# Grass2

<i> Turnbull, L. A., J. M. Levine, M. Loreau, and A. Hector. 2013. Coexistence, niches and biodiversity effects on ecosystem functioning. Ecology Letters 16:116–127. </i> 

```{r Grass2-setup, include=FALSE}

Grass2 <- models %>%
    filter(Model == "lindsay") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Grass2-plot-64}

ggplot(Grass2 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass2-remove-64, include=FALSE}

Grass2 <- Grass2 %>%
    filter(Ninitial != 64)

```

## Monoculture vs. mixture biomass

First we plot the relationship between a species' monoculture biomass and its biomass within the mixture datasets to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within each planted species richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Grass2-plot}

ggplot(Grass2)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage),
               rows = vars(Ninitial)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

## Correlations

Within each planted species richness treatment and seed addition stage, we calculate the correlation between the species' biomasses in the mixtures and their biomasses in monoculture.

```{r Grass2-correlation, echo = TRUE}

Grass2.corr <- Grass2 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Grass2.corr <- Grass2.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Grass2.corr, 
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```

## Bootstrap

To mitigate the effect of sampling in our final analysis and generate a more reliable assessment of the average correlation between the species' biomass in monoculture and their biomass in mixture, rather than only using the 64 mixtures, we bootstrap the mixture datasets (n = 2500) and calculate the Pearson's correlation for each bootstrap.

```{r Grass2-bootstrap}

Grass2.bootstrap <- Grass2 %>%
    group_by(Ninitial, Stage) %>%
    nest() %>%
    mutate(cors = map(.x = data,
                      .f = ~
                          {
                              bootstraps <- .x %>% rsample::bootstraps(times = 2500)
                              
                              bootstraps <- bootstraps %>%
                                  mutate(r = map_dbl(.x = splits,
                                                     .f = ~ 
                                                         {
                                                             .x <- as.data.frame(.x)
                                                             cor(.x$Biomass, .x$monoculture.biomass)
                                                         }))
                          })) %>%
    unnest(cors)

ggplot(Grass2.bootstrap) +
    geom_histogram(aes(x = r)) +
    facet_grid(rows = vars(Ninitial),
               cols = vars(Stage), 
               scales = "free") +
    labs(x = "Pearson's correlation (r)",
         y = "Frequency") +
    theme_bw()

```

Based on the bootstrapping process, we now have a more complete picture of the distribution of Pearson's correlation of the species' mixture biomass to their monoculture biomass. We use this distribution of Pearson's correlations in the main analysis by relating each bootstrapped correlation to an estimate of the relationship between realized Shannon diversity and biomass, drawn from the posterior distribution of the BRMS models.


****************************************************************************************************
# Grass3

<i> May, F., V. Grimm, and F. Jeltsch. 2009. Reversed effects of grazing on plant diversity: The role of below-ground competition and size symmetry. Oikos 118:1830–1843. </i> 

```{r Grass3-setup, include=FALSE}

Grass3 <- models %>%
    filter(Model == "IBC_grass.NDD") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Grass3-plot-64}

ggplot(Grass3 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Grass3-remove-64, include=FALSE}

Grass3 <- Grass3 %>%
    filter(Ninitial != 64)

```

## Monoculture vs. mixture biomass

First we plot the relationship between a species' monoculture biomass and its biomass within the mixture datasets to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within each planted species richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Grass3-plot}

ggplot(Grass3)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage),
               rows = vars(Ninitial)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

## Correlations

Within each planted species richness treatment and seed addition stage, we calculate the correlation between the species' biomasses in the mixtures and their biomasses in monoculture.

```{r Grass3-correlation, echo = TRUE}

Grass3.corr <- Grass3 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Grass3.corr <- Grass3.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Grass3.corr, 
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```

## Bootstrap

To mitigate the effect of sampling in our final analysis and generate a more reliable assessment of the average correlation between the species' biomass in monoculture and their biomass in mixture, rather than only using the 64 mixtures, we bootstrap the mixture datasets (n = 2500) and calculate the Pearson's correlation for each bootstrap.

```{r Grass3-bootstrap}

Grass3.bootstrap <- Grass3 %>%
    group_by(Ninitial, Stage) %>%
    nest() %>%
    mutate(cors = map(.x = data,
                      .f = ~
                          {
                              bootstraps <- .x %>% rsample::bootstraps(times = 2500)
                              
                              bootstraps <- bootstraps %>%
                                  mutate(r = map_dbl(.x = splits,
                                                     .f = ~ 
                                                         {
                                                             .x <- as.data.frame(.x)
                                                             cor(.x$Biomass, .x$monoculture.biomass)
                                                         }))
                          })) %>%
    unnest(cors)

ggplot(Grass3.bootstrap) +
    geom_histogram(aes(x = r)) +
    facet_grid(rows = vars(Ninitial),
               cols = vars(Stage), 
               scales = "free") +
    labs(x = "Pearson's correlation (r)",
         y = "Frequency") +
    theme_bw()

```

Based on the bootstrapping process, we now have a more complete picture of the distribution of Pearson's correlation of the species' mixture biomass to their monoculture biomass. We use this distribution of Pearson's correlations in the main analysis by relating each bootstrapped correlation to an estimate of the relationship between realized Shannon diversity and biomass, drawn from the posterior distribution of the BRMS models.


****************************************************************************************************
# Forest1

<i> Rüger, N., R. Condit, D. H. Dent, S. J. DeWalt, S. P. Hubbell, J. W. Lichstein, O. R. Lopez, C. Wirth, and C. E. Farrior. 2020. Demographic trade-offs predict tropical forest dynamics. Science 368:165–168. </i> 

```{r Forest1-setup, include=FALSE}

Forest1 <- models %>%
    filter(Model == "PPA") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Forest1-plot-64}

ggplot(Forest1 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Forest1-remove-64, include=FALSE}

Forest1 <- Forest1 %>%
    filter(Ninitial != 64)

```

## Monoculture vs. mixture biomass

First we plot the relationship between a species' monoculture biomass and its biomass within the mixture datasets to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within each planted species richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Forest1-plot}

ggplot(Forest1)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage),
               rows = vars(Ninitial)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

## Correlations

Within each planted species richness treatment and seed addition stage, we calculate the correlation between the species' biomasses in the mixtures and their biomasses in monoculture.

```{r Forest1-correlation, echo = TRUE}

Forest1.corr <- Forest1 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Forest1.corr <- Forest1.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Forest1.corr, 
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```

## Bootstrap

To mitigate the effect of sampling in our final analysis and generate a more reliable assessment of the average correlation between the species' biomass in monoculture and their biomass in mixture, rather than only using the 64 mixtures, we bootstrap the mixture datasets (n = 2500) and calculate the Pearson's correlation for each bootstrap.

```{r Forest1-bootstrap}

Forest1.bootstrap <- Forest1 %>%
    group_by(Ninitial, Stage) %>%
    nest() %>%
    mutate(cors = map(.x = data,
                      .f = ~
                          {
                              bootstraps <- .x %>% rsample::bootstraps(times = 2500)
                              
                              bootstraps <- bootstraps %>%
                                  mutate(r = map_dbl(.x = splits,
                                                     .f = ~ 
                                                         {
                                                             .x <- as.data.frame(.x)
                                                             cor(.x$Biomass, .x$monoculture.biomass)
                                                         }))
                          })) %>%
    unnest(cors)

ggplot(Forest1.bootstrap) +
    geom_histogram(aes(x = r)) +
    facet_grid(rows = vars(Ninitial),
               cols = vars(Stage), 
               scales = "free") +
    labs(x = "Pearson's correlation (r)",
         y = "Frequency") +
    theme_bw()

```

Based on the bootstrapping process, we now have a more complete picture of the distribution of Pearson's correlation of the species' mixture biomass to their monoculture biomass. We use this distribution of Pearson's correlations in the main analysis by relating each bootstrapped correlation to an estimate of the relationship between realized Shannon diversity and biomass, drawn from the posterior distribution of the BRMS models.


****************************************************************************************************
# Forest2

<i> Maréchaux, I., and J. Chave. 2017. An individual-based forest model to jointly simulate carbon and tree diversity in Amazonia: description and applications. Ecological Monographs. </i> 

```{r Forest2-setup, include=FALSE}

Forest2 <- models %>%
    filter(Model == "TROLL") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Forest2-plot-64}

ggplot(Forest2 %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Forest2-remove-64, include=FALSE}

Forest2 <- Forest2 %>%
    filter(Ninitial != 64)

```

## Monoculture vs. mixture biomass

First we plot the relationship between a species' monoculture biomass and its biomass within the mixture datasets to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within each planted species richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Forest2-plot}

ggplot(Forest2)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage),
               rows = vars(Ninitial)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

## Correlations

Within each planted species richness treatment and seed addition stage, we calculate the correlation between the species' biomasses in the mixtures and their biomasses in monoculture.

```{r Forest2-correlation, echo = TRUE}

Forest2.corr <- Forest2 %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Forest2.corr <- Forest2.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Forest2.corr, 
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```

## Bootstrap

To mitigate the effect of sampling in our final analysis and generate a more reliable assessment of the average correlation between the species' biomass in monoculture and their biomass in mixture, rather than only using the 64 mixtures, we bootstrap the mixture datasets (n = 2500) and calculate the Pearson's correlation for each bootstrap.

```{r Forest2-bootstrap}

Forest2.bootstrap <- Forest2 %>%
    group_by(Ninitial, Stage) %>%
    nest() %>%
    mutate(cors = map(.x = data,
                      .f = ~
                          {
                              bootstraps <- .x %>% rsample::bootstraps(times = 2500)
                              
                              bootstraps <- bootstraps %>%
                                  mutate(r = map_dbl(.x = splits,
                                                     .f = ~ 
                                                         {
                                                             .x <- as.data.frame(.x)
                                                             cor(.x$Biomass, .x$monoculture.biomass)
                                                         }))
                          })) %>%
    unnest(cors)

ggplot(Forest2.bootstrap) +
    geom_histogram(aes(x = r)) +
    facet_grid(rows = vars(Ninitial),
               cols = vars(Stage), 
               scales = "free") +
    labs(x = "Pearson's correlation (r)",
         y = "Frequency") +
    theme_bw()

```

Based on the bootstrapping process, we now have a more complete picture of the distribution of Pearson's correlation of the species' mixture biomass to their monoculture biomass. We use this distribution of Pearson's correlations in the main analysis by relating each bootstrapped correlation to an estimate of the relationship between realized Shannon diversity and biomass, drawn from the posterior distribution of the BRMS models.


****************************************************************************************************
# Dryland

<i> Reineking, B., M. Veste, C. Wissel, and A. Huth. 2006. Environmental variability and allocation trade-offs maintain species diversity in a process-based model of succulent plant communities. Ecological Modelling.</i> 

```{r Dryland-setup, include=FALSE}

Dryland <- models %>%
    filter(Model == "bjoern") %>%
    select(-Model)

```

## 64-species mixture

Initially, to demonstrate the broader competitive hierarchy within the model, we plot the biomass of all species within the 64-species treatment, during the terminal time step of both the metacommunity and isolation phases.

```{r Dryland-plot-64}

ggplot(Dryland %>% filter(Ninitial == 64))  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

```{r Dryland-remove-64, include=FALSE}

Dryland <- Dryland %>%
    filter(Ninitial != 64)

```

## Monoculture vs. mixture biomass

First we plot the relationship between a species' monoculture biomass and its biomass within the mixture datasets to visualize the intent of this part of the analysis. There are 64 mixtures (i.e. replicate plots) within each planted species richness treatment. Each color is a distinct species, and each point is a species' biomass within one mixture.

```{r Dryland-plot}

ggplot(Dryland)  +
    geom_point(aes(x = monoculture.biomass,
                   y = Biomass,
                   color = SpeciesID),
               show.legend = FALSE) +
    facet_grid(cols = vars(Stage),
               rows = vars(Ninitial)) +
    scale_color_viridis_d() +
    labs(x = "Monoculture biomass",
         y = "Biomass in mixture") +
    theme_bw()

```

## Correlations

Within each planted species richness treatment and seed addition stage, we calculate the correlation between the species' biomasses in the mixtures and their biomasses in monoculture.

```{r Dryland-correlation, echo = TRUE}

Dryland.corr <- Dryland %>%
    group_by(Ninitial, Stage) %>%
    summarise(corr = cor(Biomass, monoculture.biomass))

Dryland.corr <- Dryland.corr %>%
    pivot_wider(names_from = Stage, values_from = corr)

knitr::kable(Dryland.corr, 
             col.names = c("Planted species richness", "Metacommunity", "Isolation"), 
             align = 'r',
             format = "html", 
             digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = F)

```

## Bootstrap

To mitigate the effect of sampling in our final analysis and generate a more reliable assessment of the average correlation between the species' biomass in monoculture and their biomass in mixture, rather than only using the 64 mixtures, we bootstrap the mixture datasets (n = 2500) and calculate the Pearson's correlation for each bootstrap.

```{r Dryland-bootstrap}

Dryland.bootstrap <- Dryland %>%
    group_by(Ninitial, Stage) %>%
    nest() %>%
    mutate(cors = map(.x = data,
                      .f = ~
                          {
                              bootstraps <- .x %>% rsample::bootstraps(times = 2500)
                              
                              bootstraps <- bootstraps %>%
                                  mutate(r = map_dbl(.x = splits,
                                                     .f = ~ 
                                                         {
                                                             .x <- as.data.frame(.x)
                                                             cor(.x$Biomass, .x$monoculture.biomass)
                                                         }))
                          })) %>%
    unnest(cors)

ggplot(Dryland.bootstrap) +
    geom_histogram(aes(x = r)) +
    facet_grid(rows = vars(Ninitial),
               cols = vars(Stage), 
               scales = "free") +
    labs(x = "Pearson's correlation (r)",
         y = "Frequency") +
    theme_bw()

```

Based on the bootstrapping process, we now have a more complete picture of the distribution of Pearson's correlation of the species' mixture biomass to their monoculture biomass. We use this distribution of Pearson's correlations in the main analysis by relating each bootstrapped correlation to an estimate of the relationship between realized Shannon diversity and biomass, drawn from the posterior distribution of the BRMS models.
