---
title: "Traits (all models) Vs. log Mean biomass in metacommunity and isolation"
author: "VeronikaCeballosN"
date: "4/30/2021"
output: html_document
---

## Setup

```{r setup, include=FALSE}
options(width = 1000)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 4,
                      fig.width = 8,
                      size = "small") 

library(ggplot2)
library(viridis) 
library(ggpubr)
```

## Read models

First we read the 6 models, using the "readModels.R" script from Michael Crawford.

```{r read-models, echo=TRUE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)); setwd("../../")
base_dir          <- getwd()
scripts_dir       <- paste0(base_dir, "/R")
raw_data_dir      <- paste0(base_dir, "/data/raw")
tmp_dir           <- paste0(base_dir, "/tmp")

source(paste0(scripts_dir, "/readModels.R"))
```

### List of model parameters:

Grass 1
- abmi (monoculture biomass)
- pNi (aboveground N concentration)
- no3i (nitrogen R*)

Grass 2
- theta_i (Nitrogen uptake rate per unit plant biomass)
- v_i (volume of soil accessible to species i)

Grass 3
- Gmax (maximum resource utilization per time step)
- LMR (leaf to mass ratio)
- MaxMass (Plant's maximum size)
- MeanSpacerLength 
- SLA (specific leaf area)

Forest 1
- PC1score (component associated with fast-slow lifecycle) #wood density is the best related functional trait#plant height
- PC2score (component associated with tree stature) #LMA -leaf mass per area

Forest 2
- LMA (leaf mass per area)
- N - nmass (leaf N content per dry mass)
- P - pmass (leaf P content per dry mass)
- wsg (wood specific gravity)
- dbh_thresh - dmax (diameter at breast height threshold)
- h_lim (asymptotic height)
- a_h (parameter of the tree height -dbh allometry)
- h_realmax

Dryland
- maxSize (maximum size/size at maturity) [gC]
- pLeaf (allocation to leaf) [gC/gC]
- pRoot (allocation to root) [gC/gC]
- pStorage (allocation to storage) [gC/gC]

The next step is to arrange the data frames with the trait values and biomass for the model runs in monoculture and in 32 species mixtures (at the 100th -metacommunity- and 200th -isolation- year time point), and plot it. 
For models with >3 traits, we only plot the 2 traits that have the highest correlation with biomass, according to the random forest results obtained prior to this analysis (see Model Validation in SimNet Github repository). Otherwise, we plot combinations of 2 traits. 

The plotted variables are: trait 1 Vs. Trrait 2, with dots size according to mean biomass (log scale), and colored according to the Species ID.

For this purpose we have the following function:

```{r plotting_function, echo=TRUE}
t1_vs_t2_vs_biomass_meta_iso <- function(plotName,model,NoSpp,g_by,Biomass,trait1,trait2,xlab,ylab){
labNoSpp <- if (NoSpp == 32) "32-species mixture" else  "monoculture"
#labNoSpp <- if (NoSpp == 1) "monoculture"

      model.32 <- model %>%
      filter(Ninitial == NoSpp,
             Year %in% c(100, 200)) %>% 
             #100th year: last step of Metacommunity stage 
             #200th year: last step of Isolation stage
      group_by_at(g_by) %>%
      summarise(Biomass = mean(Biomass)) %>%
      filter(Biomass > 0) %>%
      mutate(Biomass = log(Biomass)) %>%
      mutate(Stage = recode(Stage, "assembly" = "Metacommunity", "With seed inflow" = "Metacommunity", 
                                   "disassembly" = "Isolation", "Without seed inflow" = "Isolation"))
      model.032 <- model %>%
      filter(Ninitial == NoSpp,
             Year %in% c(100, 200)) %>% 
      group_by_at(g_by) %>%
      summarise(Biomass = mean(Biomass)) %>%
      filter(Biomass == 0) %>%
      mutate(Stage = recode(Stage, "assembly" = "Metacommunity", "With seed inflow" = "Metacommunity", 
                                   "disassembly" = "Isolation", "Without seed inflow" = "Isolation"))
          
  p1 <- ggplot() +
    geom_point(data = model.32,
               aes_string(x = trait1,
                          y = trait2,
                          color = Biomass, alpha = 0.7,
                          size = Biomass),
               position=position_jitter(h=0.01, w=0.05)) +
    geom_point(data = model.032, shape = 4,
               aes_string(x = trait1,
                          y = trait2),
               position=position_jitter(h=0.01, w=0.05)) +
    facet_grid(~fct_relevel(Stage, 'Metacommunity', 'Isolation')) +
    scale_color_viridis() +
    labs(size = "Log mean biomass",
	 color = "Log mean biomass",
         x = xlab,
         y = ylab) +
    theme_bw() +
    theme(aspect.ratio = 1)
  
  filename <- paste0(plotName,".pdf")
  path <- paste0(tmp_dir,"/pdfs/")
  ggsave(filename = filename
         ,path = path
         ,plot = p1
         ,height = 12
         ,width = 19
         ,units = "cm")
  return(p1)
}
```

We proceed to use the above mentioned function for each model:

### Grass1 (Adam's model)

Grass1 has 3 variables, but here only two (pNi -aboveground tissue nitrogen concentration- and no3i -nitrogen R*-) will be plotted, because the trait 'abmi' corresponds to monoculture biomass.

```{r Grass1-traits-vs-32spp-biomass, echo=TRUE, fig.height = 6, fig.width = 18}
#source(paste0(scripts_dir, "/to_test/readAdam.R"))

adam <- models$Grass1 %>%
  ungroup() %>% # There shouldn't be groups anyways
  mutate(Biomass = scales::rescale(Biomass,
                                   to = c(0, 100)))%>%
  select(-Model, -SeedRain) 

g_by <- c("SpeciesID", "Year", "Stage", "pNi", "no3i")
lab1 <- "pNi (aboveground N concentration)"
lab2 <- "no3i (nitrogen R*)"

p1 <- t1_vs_t2_vs_biomass_meta_iso("G1mono",adam,1,g_by,"Biomass","pNi","no3i",lab1,lab2)
p2 <- t1_vs_t2_vs_biomass_meta_iso("G1mix",adam,32,g_by,"Biomass","pNi","no3i",lab1,lab2)
#ggarrange(p1, p2, ncol = 2, nrow = 1)
```

### Grass2 (Lindsay's model)

Grass2 has 2 traits: Theta_i and V_i

```{r Grass2-traits-vs-32spp-biomass, echo=TRUE, fig.height = 6, fig.width = 18}
#source(paste0(scripts_dir, "/to_test/readLindsay.R"))

lindsay <- models$Grass2%>%
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))%>%
    select(-Model, -SeedRain)

g_by <- c("SpeciesID", "Year", "Stage", "Vi", "thetai")
lab1 <- "Vi (volume of soil accessible to species i)"
lab2 <- "thetai (Nitrogen uptake rate per unit plant biomass)"

p3 <- t1_vs_t2_vs_biomass_meta_iso("G2mono",lindsay,1,g_by,"Biomass","Vi","thetai",lab1,lab2)
p4 <- t1_vs_t2_vs_biomass_meta_iso("G2mix",lindsay,32,g_by,"Biomass","Vi","thetai",lab1,lab2)
```

### Grass3 (IBC-grass)

Grass3 has 5 traits: Gmax, LMR, MaxMass, MeanSpacerLength, and SLA. After the randomForest analysis, Gmax and LMR were selected.

```{r Grass3-traits-vs-32spp-biomass, echo=TRUE, fig.height = 6, fig.width = 18}
#source(paste0(scripts_dir, "/to_test/readIBC.R"))

IBC_grass <- models$Grass3 %>%
  ungroup() %>% # There shouldn't be groups anyways
  mutate(Biomass = scales::rescale(Biomass,
                                   to = c(0, 100)))%>%
  select(-Model, -SeedRain)

g_by <- c("SpeciesID", "Year", "Stage", "LMR", "MaxMass", "Gmax", "SLA", "meanSpacerLength")
lab1 <- "Gmax (maximum resource utilization per time step)"
lab2 <- "MaxMass (Plant's maximum size)"
lab3 <- "LMR (leaf to mass ratio)"

p5 <- t1_vs_t2_vs_biomass_meta_iso("G3mono",IBC_grass,1,g_by,"Biomass","Gmax","MaxMass",lab1,lab2)
p6 <- t1_vs_t2_vs_biomass_meta_iso("G3mix",IBC_grass,32,g_by,"Biomass","Gmax","MaxMass",lab1,lab2)
```

### Forest1 (PPA)

Forest1 has 2 traits: PC1score and PC2score.

```{r Forest1-traits-vs-32spp-biomass, echo=TRUE, fig.height = 6, fig.width = 18}
#source(paste0(scripts_dir, "/to_test/readPPA.R"))

PPA <- models$Forest1 %>%
  ungroup() %>% # There shouldn't be groups anyways
  mutate(Biomass = scales::rescale(Biomass,
                                   to = c(0, 100)))%>%
  select(-Model, -SeedRain)

g_by <- c("SpeciesID", "Year", "Stage", "PC1score", "PC2score")
lab1 <- "PC1score (associated with fast-slow lifecycle)"#plant height)"
lab2 <- "PC2score (associated with tree stature)"# LMA -leaf mass per area)"

p7 <- t1_vs_t2_vs_biomass_meta_iso("F1mono",PPA,1,g_by,"Biomass","PC1score","PC2score",lab1,lab2)
p8 <- t1_vs_t2_vs_biomass_meta_iso("F1mix",PPA,32,g_by,"Biomass","PC1score","PC2score",lab1,lab2)
```

### Forest2 (TROLL)

Forest2 has 7 traits: LMA (leaf mass per area), N -nmass- (leaf N content per dry mass), P -pmass- (leaf P content per dry mass), wsg (wood specific gravity), dbh_thresh -dmax- (diameter at breast height threshold), -h_realmax- {h_lim -hmax- (asymptotic height), ah (parameter of the tree-height-dbh allometry)}. 

After the randomForest analysis, LMA and dmax were selected.

```{r Forest2-traits-vs-32spp-biomass, echo=TRUE, fig.height = 6, fig.width = 18}
#source(paste0(scripts_dir, "/to_test/readTROLL.R"))

troll <- models$Forest2 %>%
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))%>%
    select(-Model, -SeedRain) %>%
    mutate(h_realmax = hmax * dmax / (dmax + ah)) %>%
    select(-hmax, -ah)

g_by <- c("SpeciesID", "Year","Stage", "dmax", "wsg", "lma", "pmass", "nmass", "h_realmax")
lab1 <- "lma (leaf mass per area)"
lab2 <- "dmax (diameter at breast height threshold)"
lab3 <- "wsg (wood specific gravity)"
lab4 <- "pmass (leaf P content per dry mass)"

p9 <- t1_vs_t2_vs_biomass_meta_iso("F2mono",troll,1,g_by,"Biomass","dmax","wsg",lab2,lab3)
p10 <- t1_vs_t2_vs_biomass_meta_iso("F2mix",troll,32,g_by,"Biomass","lma","dmax",lab1,lab2)
```

### Dryland (Bjoern)

Dryland has 4 traits. After the randomForest analysis, maxSize and pLeaf were selected.

```{r Dryland-traits-vs-32spp-biomass, echo=TRUE, fig.height = 6, fig.width = 18}
#source(paste0(scripts_dir, "/to_test/readBjoern.R"))

bjoern <- models$bjoern %>% #Dryland %>%
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))%>%
    select(-Model, -SeedRain) 

g_by <- c("SpeciesID", "Year","Stage", "maxSize", "pLeaf", "pRoot", "pStorage")
lab1 <- "maxSize (maximum size/size at maturity) [gC]"
lab2 <- "pLeaf (allocation to leaf) [gC/gC]"
lab3 <- "pStorage (allocation to storage) [gC/gC]"

p11 <- t1_vs_t2_vs_biomass_meta_iso("Dmono",bjoern,1,g_by,"Biomass","maxSize","pStorage",lab1,lab3)
p12 <- t1_vs_t2_vs_biomass_meta_iso("Dmix",bjoern,32,g_by,"Biomass","maxSize","pLeaf",lab1,lab2)
```
## All models

4 conditions: metacommunity vs. isolation, monoculture vs. mixture

- Grass 1:

Biomass is highest with high values of aboveground N concentration (pNi) and nitrogen R* (no3i). The trend seems to be mantained in all four conditions (metacommunity vs. isolation, monoculture vs. mixture); however, the seed inflow seems to play an important role in the survival of some species, whose biomass decreases in isolation, and in some cases they do not survive at the end of the isolation phase. 

- Grass 2:

The species that seem to dominate in mixture are not necessarily those with the highest biomass accummulation in monoculture. Given all four conditions, the highest biomass values are obtained by the species that have accessibility to a high volume of soil (Vi) and low nitrogen uptake rate per unit plant biomass (thetai). Species growing in the mixture seem to depend more on the seed inflow to keep their growth, than those in monoculture. 

- Grass 3:

In monoculture, the species with the lowest Gmax (maximum resource utilization per time step) grow the most, regardless of their leaf to mass ratio (LMR). This trend is less strong in mixture, where the growth of species is similar at different LMR and Gmax values. The seed inflow maintains the survival of species that have >20 Gmax values and grow in mixture.

- Forest 1:

Tall species (PC2score) with a slow lifecycle (PC1score) accumulate the highest biomass in all conditions. Some species do not survive in mixture (particularly those that are short and have a slow life cycle); seed inflow keeps survival of most species in mixture (see that the amount of species that do not survive -marked with an "x"- in isolation  exceeds those that do).

- Forest 2:

Species with the highest values of dmax (diameter at breast height threshold) and wsg (wood specific gravity) have the highest biomass in monoculture. In mixture, another trait different to wsg has an impact on the biomass: lma (leaf mass per area); species with the lowest values of lma grow the most, at any dmax. Seed inflow only has an effect in mixture, where the biomass accumulation seems to diminish by the end of the isolation period and some species do not survive.

- Dryland:

In monoculture, species with an average allocation to storage (pStorage) (as long as their size at maturity -maxSize surpasses a low threshold). In mixture, pSotarge does not influence biomass accumulation as much as pLeaf does (see random Forest results for this model in Crawford's Mechanisms Rmd files). For the species growing in mixture, the clearest trend is observed after the isolation phase, where the species with the highest biomass are those with medium sized allocation to leaves (pLeaf) and low maximum size at maturity. Similar to forest 2, the species in mixture are more affected to seed inflow than those in monoculture; by the end of the isolation phase the biomass accumulation seems to diminish and a few species do not survive.

```{r traits-vs-32spp-biomass, echo=TRUE, fig.height = 33, fig.width = 20}
ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12, 
         labels = c("Grass 1 monoculture","Grass 1 mixture","Grass 2 monoculture","Grass 2 mixture","Grass 3 monoculture","Grass 3 mixture","Forest 1 monoculture","Forest 1 mixture","Forest 2 monoculture","Forest 2 mixture","Dryland monoculture","Dryland mixture"),
         ncol = 2, nrow = 6)

```
