---
title: "Random Forest Analysis: Traits (all models) Vs. log Mean biomass"
author: "VeronikaCeballosN"
date: "4/30/2021"
output: html_document
---

## Setup

```{r setup, include=FALSE}

options(width = 1000)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      #fig.height = 4,
                      #fig.width = 8,
                      size = "small") 

library(data.table)
library(tidyverse)
library(ggthemes)
library(gganimate)
library(viridis) 
library(cowplot)
library(scales)
library(plotly)
library(randomForest)
library(pdp)

```

## Read models

First we read the 6 models, using the "readModels.R" script from Michael Crawford.

```{r read-models, echo=TRUE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)); setwd("../../")
base_dir          <- getwd()
scripts_dir       <- paste0(base_dir, "/R")
tmp_dir           <- paste0(base_dir, "/tmp")
raw_data_dir      <- paste0(base_dir, "/data/raw")

source(paste0(scripts_dir, "/readModels.R"))
```

```{r randomForest_4conditions, echo=TRUE}
meta = seq(80, 100)
iso = seq(180, 200)

randomForest_4conditions <- function(modelName,model,NoSpp,stage){
        model <- model %>%
            filter(Ninitial == NoSpp,
                   Year %in% stage) %>%
            mutate(id = row_number()) %>%
            mutate_if(is.character, as.factor) %>%
            select(-Productivity,-SpeciesID, -Ninitial, -Stage, -Rep) 
        
        train <- model %>% sample_frac(.70)
        test <- anti_join(model, train, by = 'id')
        
        train <- train %>% select(-id)
        test <- test %>% select(-id)
        
        rf <- randomForest(data = as.data.frame(train),
                                        Biomass ~ .,
                                        importance = TRUE)
        
        pred <- data.frame(pred = predict(rf, test))
        title = paste("correlation: ", round(cor(pred, test$Biomass)[[1]], 2),
                      "  |  mean squared error: ", round(mean(rf$mse), 2),
                      "  |  R-squared: ", round(mean(rf$rsq), 2),
                      sep = "")
        pdf(paste0(tmp_dir,"/randomForest/",modelName,".pdf"))
        varImpPlot(rf, main = title)
        #dev.set(dev.next())
        while (!is.null(dev.list()))  dev.off()
}
```

### Grass1 (Adam's model)

Grass1 has 3 variables, but here only two (pNi -aboveground tissue nitrogen concentration- and no3i -nitrogen R*-) will be plotted, because the trait 'abmi' corresponds to monoculture biomass.

```{r Grass1-randomForest1, echo=TRUE, fig.height = 6, fig.width = 12}
adam <- adam %>%
    select(-Model, -SeedRain) %>%
    mutate(SpeciesID = as.factor(SpeciesID))

adam_traits <- adam_traits %>%
    mutate(SpeciesID = as.factor(SpeciesID))

adam <- inner_join(adam, adam_traits, by = c("SpeciesID"))

G1C1 <- randomForest_4conditions("G1C1",adam,1,meta)
G1C2 <- randomForest_4conditions("G1C2",adam,1,iso)
G1C3 <- randomForest_4conditions("G1C3",adam,32,meta)
G1C4 <- randomForest_4conditions("G1C4",adam,32,iso)
```

### Grass2 (Lindsay's model)

Grass2 has 2 traits: Theta_i and V_i

```{r Grass2-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
lindsay <- lindsay %>%
    select(-Model, -SeedRain) %>%
    mutate(SpeciesID = as.factor(SpeciesID))

lindsay_traits <- lindsay_traits %>%
    mutate(SpeciesID = as.factor(SpeciesID))

lindsay <- inner_join(lindsay, lindsay_traits, by = c("SpeciesID"))

G2C1 <- randomForest_4conditions("G2C1",lindsay,1,meta)
G2C2 <- randomForest_4conditions("G2C2",lindsay,1,iso)
G2C3 <- randomForest_4conditions("G2C3",lindsay,32,meta)
G2C4 <- randomForest_4conditions("G2C4",lindsay,32,iso)
```

### Grass3 (IBC-grass)

Grass3 has 5 traits: Gmax, LMR, MaxMass, MeanSpacerLength, and SLA. After the randomForest analysis, Gmax and LMR were selected.

```{r Grass3-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
IBC_grass <- models$Grass3 %>%
    select(-Model, -SeedRain) %>%
    mutate(SpeciesID = as.factor(SpeciesID))

IBC_grass_traits <- IBC_grass_traits %>%
    mutate(SpeciesID = as.factor(SpeciesID))

IBC_grass <- inner_join(IBC_grass, IBC_grass_traits, by = c("SpeciesID"))

G3C1 <- randomForest_4conditions("G3C1",IBC_grass,1,meta)
G3C2 <- randomForest_4conditions("G3C2",IBC_grass,1,iso)
G3C3 <- randomForest_4conditions("G3C3",IBC_grass,32,meta)
G3C4 <- randomForest_4conditions("G3C4",IBC_grass,32,iso)
```

### Forest1 (PPA)

Forest1 has 2 traits: PC1score and PC2score.

```{r Forest1-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
PPA <- models$Forest1 %>%
    select(-Model, -SeedRain)

F1C1 <- randomForest_4conditions("F1C1",PPA,1,meta)
F1C2 <- randomForest_4conditions("F1C2",PPA,1,iso)
F1C3 <- randomForest_4conditions("F1C3",PPA,32,meta)
F1C4 <- randomForest_4conditions("F1C4",PPA,32,iso)
```

### Forest2 (TROLL)

Forest2 has 7 traits: LMA (leaf mass per area), N -nmass- (leaf N content per dry mass), P -pmass- (leaf P content per dry mass), wsg (wood specific gravity), dbh$_{thresh}$ -dmax- (diameter at breast height threshold), -h_realmax- (h$_{lim}$ -hmax- (asymptotic height), ah (parameter of the tree-height-dbh allometry)). 

After the randomForest analysis, LMA and dmax were selected.

```{r Forest2-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
troll <- troll %>%
    select(-Model, -SeedRain) %>%
    mutate(SpeciesID = as.factor(SpeciesID))

troll_traits <- troll_traits %>%
    mutate(SpeciesID = as.factor(SpeciesID))

troll <- inner_join(troll, troll_traits, by = c("SpeciesID"))

troll <- troll %>%
    mutate(h_realmax = hmax * dmax / (dmax + ah)) %>%
    select(-hmax, -ah)

F2C1 <- randomForest_4conditions("F2C1",troll,1,meta)
F2C2 <- randomForest_4conditions("F2C2",troll,1,iso)
F2C3 <- randomForest_4conditions("F2C3",troll,32,meta)
F2C4 <- randomForest_4conditions("F2C4",troll,32,iso)
```

### Dryland (Bjoern)

Dryland has 4 traits. After the randomForest analysis, maxSize and pLeaf were selected.

```{r Dryland-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
bjoern <- models$bjoern %>%#Dryland %>% 
    select(-Model, -SeedRain) %>%
    mutate(SpeciesID = as.factor(SpeciesID))

bjoern_traits <- model_traits[[6]] %>%
    mutate(SpeciesID = as.factor(SpeciesID))

bjoern <- inner_join(bjoern, bjoern_traits, by = c("SpeciesID"))

DC1 <- randomForest_4conditions("DC1",bjoern,1,meta)
DC2 <- randomForest_4conditions("DC2",bjoern,1,iso)
DC3 <- randomForest_4conditions("DC3",bjoern,32,meta)
DC4 <- randomForest_4conditions("DC4",bjoern,32,iso)
```
