---
title: "Random Forest Analysis: Traits (all models) Vs. log Mean biomass"
author: "VeronikaCeballosN"
date: "4/30/2021"
output: html_document
---

## Setup

```{r setup, include=FALSE}

options(width = 1000)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      #fig.height = 4,
                      #fig.width = 8,
                      size = "small") 
```

## Read models

First we read the 6 models, using the "readModels.R" script from Michael Crawford.

```{r read-models, echo=TRUE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)); setwd("../../")
base_dir          <- getwd()
scripts_dir       <- paste0(base_dir, "/R")
tmp_dir           <- paste0(base_dir, "/tmp")
raw_data_dir      <- paste0(base_dir, "/data/raw")

source(paste0(scripts_dir, "/readModels.R"))

source(paste0(scripts_dir, "/to_test/randomForest_4conditions.R"))
```

### Grass1 (Adam's model)

Grass1 has 3 variables, but here only two (pNi -aboveground tissue nitrogen concentration- and no3i -nitrogen R*-) will be plotted, because the trait 'abmi' corresponds to monoculture biomass.

```{r Grass1-randomForest1, echo=TRUE, fig.height = 6, fig.width = 12}
adam <- models$Grass1 %>%
  ungroup() %>% # There shouldn't be groups anyways
  mutate(Biomass = scales::rescale(Biomass,
                                   to = c(0, 100)))%>%
  select(-Model, -SeedRain) 

G1C1 <- randomForest_4conditions("G1C1",adam,1,meta)
G1C2 <- randomForest_4conditions("G1C2",adam,1,iso)
G1C3 <- randomForest_4conditions("G1C3",adam,32,meta)
G1C4 <- randomForest_4conditions("G1C4",adam,32,iso)
```

### Grass2 (Lindsay's model)

Grass2 has 2 traits: Theta_i and V_i

```{r Grass2-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
lindsay <- models$Grass2%>%
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))%>%
    select(-Model, -SeedRain) 

G2C1 <- randomForest_4conditions("G2C1",lindsay,1,meta)
G2C2 <- randomForest_4conditions("G2C2",lindsay,1,iso)
G2C3 <- randomForest_4conditions("G2C3",lindsay,32,meta)
G2C4 <- randomForest_4conditions("G2C4",lindsay,32,iso)
```

### Grass3 (IBC-grass)

Grass3 has 5 traits: Gmax, LMR, MaxMass, MeanSpacerLength, and SLA. After the randomForest analysis, Gmax and LMR were selected.

```{r Grass3-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
IBC_grass <- models$Grass3 %>%
  ungroup() %>% # There shouldn't be groups anyways
  mutate(Biomass = scales::rescale(Biomass,
                                   to = c(0, 100)))%>%
  select(-Model, -SeedRain) %>%
  select(Rep, Ninitial, SpeciesID, Year, Stage, Productivity, Biomass, LMR, MaxMass, Gmax, SLA, meanSpacerLength)
 

G3C1 <- randomForest_4conditions("G3C1",IBC_grass,1,meta)
G3C2 <- randomForest_4conditions("G3C2",IBC_grass,1,iso)
G3C3 <- randomForest_4conditions("G3C3",IBC_grass,32,meta)
G3C4 <- randomForest_4conditions("G3C4",IBC_grass,32,iso)
```

### Forest1 (PPA)

Forest1 has 2 traits: PC1score and PC2score.

```{r Forest1-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
PPA <- models$Forest1 %>%
  ungroup() %>% # There shouldn't be groups anyways
  mutate(Biomass = scales::rescale(Biomass,
                                   to = c(0, 100)))%>%
  select(-Model, -SeedRain)

F1C1 <- randomForest_4conditions("F1C1",PPA,1,meta)
F1C2 <- randomForest_4conditions("F1C2",PPA,1,iso)
F1C3 <- randomForest_4conditions("F1C3",PPA,32,meta)
F1C4 <- randomForest_4conditions("F1C4",PPA,32,iso)
```

### Forest2 (TROLL)

Forest2 has 7 traits: LMA (leaf mass per area), N -nmass- (leaf N content per dry mass), P -pmass- (leaf P content per dry mass), wsg (wood specific gravity), dbh$_{thresh}$ -dmax- (diameter at breast height threshold), -h_realmax- (h$_{lim}$ -hmax- (asymptotic height), ah (parameter of the tree-height-dbh allometry)). 

After the randomForest analysis, LMA and dmax were selected.

```{r Forest2-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
troll <- models$Forest2 %>%
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))%>%
    select(-Model, -SeedRain) %>%
    mutate(h_realmax = hmax * dmax / (dmax + ah)) %>%
    select(-hmax, -ah)

F2C1 <- randomForest_4conditions("F2C1",troll,1,meta)
F2C2 <- randomForest_4conditions("F2C2",troll,1,iso)
F2C3 <- randomForest_4conditions("F2C3",troll,32,meta)
F2C4 <- randomForest_4conditions("F2C4",troll,32,iso)
```

### Dryland (Bjoern)

Dryland has 4 traits. After the randomForest analysis, maxSize and pLeaf were selected.

```{r Dryland-randomForest, echo=TRUE, fig.height = 6, fig.width = 12}
bjoern <- models$bjoern %>% 
    ungroup() %>% # There shouldn't be groups anyways
    mutate(Biomass = scales::rescale(Biomass,
                                     to = c(0, 100)))%>%
    select(-Model, -SeedRain) 

DC1 <- randomForest_4conditions("DC1",bjoern,1,meta)
DC2 <- randomForest_4conditions("DC2",bjoern,1,iso)
DC3 <- randomForest_4conditions("DC3",bjoern,32,meta)
DC4 <- randomForest_4conditions("DC4",bjoern,32,iso)
```
```{r}
model_list <- c(G1C1,G1C2,G1C3,G1C4)
capture.output(for i in model_list print(i$importance),
               file = (paste0(tmp_dir,"/randomForest/importance.txt")))

```